---
title: "p8105_hw6_yl5828"
author: "Leah Li"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
editor_options: 
  chunk_output_type: console
---


## Problem 0

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(patchwork)
library(rvest)
library(broom)
library(janitor)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r read in data, message=FALSE}
url <- "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicide_data <- read_csv(url)
```


```{r data cleaning}
homicide_df <- homicide_data %>%
  mutate(city_state = paste(city, state, sep = ", ")) %>%
  mutate(resolved = as.numeric(disposition == "Closed by arrest")) %>%
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  filter(!is.na(victim_age)) %>%
  mutate(
    victim_sex = factor(victim_sex),
    victim_race = factor(victim_race)
  )
```



```{r logistic regression for Baltimore}
baltimore_df <- homicide_df %>%
  filter(city_state == "Baltimore, MD")

baltimore_glm <- glm(
  resolved ~ victim_age + victim_sex + victim_race,
  data = baltimore_df,
  family = binomial()
)

# Tidy the results
baltimore_results <- baltimore_glm %>%
  tidy() %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>%
  filter(term == "victim_sexMale") %>%
  select(term, OR, CI_lower, CI_upper)
```

Baltimore, MD Results:  

For Baltimore, the adjusted odds ratio for solving homicides comparing male victims to female victims is
`r round(baltimore_results$OR, 3)` with 95% CI: (**`r round(baltimore_results$CI_lower, 3)`**, **`r round(baltimore_results$CI_upper, 3)`**).



```{r Analysis for all cities}
city_results <- homicide_df %>%
  nest(data = -city_state) %>%
  mutate(
    model = map(data, ~glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())),
    tidied = map(model, tidy)) %>%
  unnest(tidied) %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>%
  select(city_state, OR, CI_lower, CI_upper) %>%
  arrange(OR)

city_results
```


```{r Create plot}
plot <- city_results %>%
  mutate(city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y = OR)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.3) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", alpha = 0.7) +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides: Male vs Female Victims",
    subtitle = "Adjusted for victim age and race across 47 U.S. cities",
    x = "City",
    y = "Odds Ratio (95% CI)") +
  theme_minimal()

plot
```


In the plot, most cities are clustered close to 1 and most 95% CIs cross the vertical line at 1, suggesting little evidence of large, consistent differences in clearance rates by victim sex within most cities. A few cities have ORs noticeably above 1, hinting that male-victim homicides might be more likely to be solved there, but their CIs are wide, so those estimates are fairly imprecise. Overall, there isn’t a strong pattern by city.



## Problem 2

```{r load packages}
library(p8105.datasets)
data("weather_df")
```


```{r calculation}
set.seed(1)

n_bootstrap <- 5000

boot_results <- 
  tibble(rep = 1:n_bootstrap) %>%
  mutate(
    sample = map(rep, ~ sample_n(weather_df, size = nrow(weather_df), replace = TRUE)),
    model = map(sample, ~ lm(tmax ~ tmin + prcp, data = .x)),
    glance_output = map(model, glance),
    tidy_output = map(model, tidy),
    r_squared = map_dbl(glance_output, "r.squared"),
    beta1 = map_dbl(tidy_output, ~ .x$estimate[.x$term == "tmin"]),
    beta2 = map_dbl(tidy_output, ~ .x$estimate[.x$term == "prcp"]),
    beta_ratio = beta1 / beta2
  ) %>%
  select(rep, r_squared, beta_ratio)

# Calculate 95% confidence intervals
ci_r_squared <- quantile(boot_results$r_squared, c(0.025, 0.975))
ci_beta_ratio <- quantile(boot_results$beta_ratio, c(0.025, 0.975))

ci_r_squared
ci_beta_ratio
```


```{r distribution plots}
# Create plot for R-squared distribution
plot_r_squared <- boot_results %>%
  ggplot(aes(x = r_squared)) +
  geom_density(fill = "purple", alpha=0.5) +
  labs(
    title = "Bootstrap Distribution of R²",
    x = "R²",
    y = "Count"
  ) +
  theme_minimal()

# Create plot for beta ratio distribution
plot_beta_ratio <- boot_results %>%
  ggplot(aes(x = beta_ratio)) +
  geom_density(bins = 40, fill = "blue", alpha=0.5)  +
  labs(
    title = "Bootstrap Distribution of β1 / β2",
    x = "β1 / β2 (tmin coefficient / prcp coefficient)",
    y = "Count"
  ) +
  theme_minimal()

plot_r_squared
plot_beta_ratio
```

From the plots, we could observe that the bootstrap distribution of R² is approximately normal with a slight left skew, centered around 0.94, which indicates the model consistently explains a high proportion of variance in the data. In contrast, the distribution of the ratio β1 / β2 is heavily left-skewed with a significant tail extending toward more negative values; this asymmetry suggests instability in the estimate, likely caused by the denominator which is β2 the precipitation coefficient, being very close to zero in certain bootstrap samples, inflating the ratio.



## Problem 3

```{r import dataset}
birthweight_raw = read_csv("data/birthweight.csv") |> 
  clean_names()
```





